---
title: "Election Timing during an Ongoing Crisis"
# subtitle: "Are they worth it?"
author: "William Zhu"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: pdf_document
header-includes:
    - \usepackage{setspace}
    - \doublespacing
    # - \usepackage{lineno}
    # - \linenumbers
# bibliography: references.bib
csl: american-journal-of-political-science.csl
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(stargazer)
```

<!-- ## Motivation -->

<!-- This is a section outlining how I arrived at this research proposal. While it is not part of the proposal proper, it provides a way for me to discuss some of the more mundane and practical reasons for pursuing this research. -->

<!-- When I was an undergrad, I took an interest in election timing in parliamentary democracies. I thought about doing a paper detailing the effect of fixed-term parliament legislation at both federal and provincial level on the turnout rate, but I am not sure the time is ripe for this since it may take quite a while for any (de)mobilizational effect to materialize. Most research on election timing makes heavy use of formal model, which was too sophisticated for me and sort of still is. In addition, empirical methods studying observational data aren't done a lot due to the issue of endogeneity. As a result, existing empirical studies about the effect of flexible election timing requires surveying voters, which is why I put this research topic on the back-burner and save it for later. -->

<!-- However, I just realized that COVID-19 may be a natural experiment that allows us to test some hypotheses about certain aspects of election timing empirically. Canada might offers an interesting motivating example. Both the provincial premiers of New Brunswick and British Columbia were leading minority governments before calling the snap election in the fall of 2020 to address pandemic needs, and their decision paid off as their parties won the majority of seats in their respective provincial legislatures. However, when Prime Minister Justin Trudeau called a snap election in the fall of 2021 for ostensibly the same reason, his Liberal Party's seat share dropped even further. This suggests that for the incumbents, the electoral benefit of calling snap elections in response to COVID-19 might taper off as we become further removed from the onset of the pandemic. Of course, this is a rough idea. My reason for pursuing this research comes from my interest in election timing, but I also hope to attract a broader audience by drawing on other relevant literature. Any feedback on the feasibility of the idea or suggestions to modify it are most appreciated. -->

## Introduction

A Crisis exogenous to the political-economic system, though scary at first, presents incumbents with a perfect chance to showcase their leadership without the risk of shouldering too much blame since they didn't start the crisis, thus boosting their popularity. Naturally, incumbents in democracies would attempt to translate such popularity into electoral advantage. Under the parliamentary system, one way for the prime minister (PM) to do so is to call a snap election to cash in his/her party's high approval rate, thereby allowing the party (and perhaps its coalition partners) to cruise to a poll victory.

Of course, the outcome of such political calculus remain uncertain when voters think the PM calls an election solely to make electoral gains. In the fall of 2021, when Canadian PM Justin Trudeau was leading the poll, he called a snap election ostensibly to make the voters decide which party will lead them to "finish the fight against covid-19".[^1] Other major parties see it as electoral opportunism. In the end, it did not turn out so well for Trudeau and his Liberal Party which saw its vote share fall even further from the last election. Though the Liberal Party managed to grab a few seats, it remained stuck in the minority government limbo as it did before the snap election. It seems that taking advantage of a crisis for electoral gains is not such a straightforward affair.

[^1]: See the Washington Post report by @coletta2021

This puzzle immediately raises two questions:

1.  What makes PMs more likely to call snap elections in response to exogenous crises?
2.  Does calling snap elections in response to exogenous crises decrease the support for PM's party?
3.  When do these snap elections reward PM's party?

Answers to these questions may also interest scholars and political advisors working on politicians serving fixed terms of offices as to when exogenous shocks stop paying electoral dividends since snap elections represent hard tests for observing the efficacy of turning approval rates into seats/margins.

## Brief Literature Review

The predominant theory on snap elections, a.k.a. opportunistic election timing, is built on the theory that voters retrospectively assess the performance of the incumbents [@fiorina1981; @key1966], especially in terms of economic indicators [@barreiro2007]. Incumbents, also operating under this assumption, try to call elections when the economy is doing [@ito1988; @chowdhury1993] well or manipulate the economy [@kayser2005] before the election to make electoral gains. @beckman2020 find that PMs who are allowed to call elections are less likely to manipulate economy. Systematic studies of calling snap elections for non-economic reasons are few and far between[@morgan-jones2018; @mcclean2021] since most of the time they are a result of the threat of snap elections which usually suffice to maintain party discipline without actually following through with it [@roy2012]. The author is not aware of any studies concerning snap elections called in response to exogenous shocks (natural disasters, pandemic) or endogenous ones (wars) possibly because these events are so idiosyncratic.

Electorate, perceiving such attempts as opportunistic, may accordingly punish the incumbents [@smith2004], but whether it is harsh enough to remains debatable [@blais2004; @schleiter2016; @schleiter2018; @daoust2021] @roy2012 simulated an election where incumbents gain an advantage only when they call an election immediately after they received positive coverage. It suggests that in the case of exogenous shocks, the incumbents would reap the most electoral advantage when they received the most amount of positive coverage for managing the exogenous shocks. Nevertheless, the incumbent still faces uncertainty as to when the approval rate would peak and how it would last. It is unclear whether opportunistic incumbents would wait for better coverage to boost their popularity, take advantage of it as soon as possible or somewhere in between.

## Research Questions

1.  What factors make PMs more likely to call snap elections in response to COVID-19?
2.  Do snap elections in response to COVID-19 pay off?

## Hypotheses

1.  The following factors make PMs more likely to call snap elections by increasing the electoral dividends of the snap elections:
    a.  Popularity indicators: Rising approval rate, better media coverage;
    b.  Crisis management indicators: Falling COVID-19 death rate/Rising vaccination rate;
    c.  Economic indicators: GDP growth/interest rate/inflation.
2.  The following factors make PMs less likely to call snap elections:
    a.  imminent expiration date of parliamentary term: capturing electoral advantage without seeming opportunistic
    b.  High proportion of seats in the parliaments: little need to risk any loss
3.  The following factors decrease the electoral dividends of the snap elections:
    a.  Voters' concern about incumbent opportunism;
    b.  Voters' concern that elections are inappropriate due to the exogenous shock;

## Empirical Strategy

I propose to use COVID-19 as a natural experiment since it is an exogenous shock to all countries around roughly the same time and present a similar set of challenges to all countries. No countries dodged the scourge of COVID-19. I am hoping that important variables mentioned in Hypotheses are random enough. I wonder if it is okay to use data from both state and national elections in democratic countries to test the claim.

1.  Most variables mentioned in Hypothesis 1 would be available through public opinion polls and statistical bureaux in those countries. Although I am not sure if poll data are comparable across countries. Because of this there may be a good argument to restrict the range of data to European elections/Canadian federal and provincial elections/Australian federal and state elections.
2.  Identifying snap elections in response to COVID-19. One could go through the news reports to see if commentators think they fit the bill. There may be other more objective methods like measuring the length of time between the last election and the time they called the election. Beyond that the rest of the variables mentioned in Hypothesis 2 are easy to come by by consulting the legislatures' official records on their websites.
3.  Hypotheses 1 and 2 together answer RQ 1 which concerns the likelihood of calling snap elections. I am hoping that logistic/probit regression should do the trick of estimating the effect of all the factors mentioned in the two hypotheses.
4.  Hypothesis 3 concerns the extent to which voters punish electoral opportunism, which I will measure by calculating the difference between the voting intention poll on the date the election was called and the actual vote share. I think it suffices to test whether the difference is significant enough, but I am going to think harder about how to expand on this hypothesis and RQ2.

\newpage

## Bibliography

<div id="refs"></div>

\newpage

# Appendix 

## Data Collection

To begin with, we consider elections held between 2020 and now in countries which, at the time of the election, were graded as "liberal democracies" or "electoral democracies" by V-Dem.

```{r vdemdata, echo=FALSE, results='asis'}
library(vdemdata)

vdem_2020_now <- filter(vdem, year > 2019)
identifier_list <- c("country_name", "country_id", "country_text_id")
vdem_high_level_indices_list <- c("v2x_polyarchy", "v2x_libdem", "v2x_partipdem", "v2x_delibdem", "v2x_egaldem")
vdem_regime_of_the_world <- c("v2x_regime", "v2x_regime_amb")
pt_est_dems <- filter(vdem_2020_now, v2x_regime > 1)
pt_est_dems_names <- distinct(select(pt_est_dems, country_name))
vdem_2020_now_essential <- select(pt_est_dems, one_of(c("year",identifier_list, vdem_high_level_indices_list, vdem_regime_of_the_world)  ) )

# ci_est_dems <- filter(vdem_2020_now, v2x_regime_amb > 3)
# ci_est_dems_names <- distinct(select(ci_est_dems, country_name))

knitr::kable(pt_est_dems_names)
```



